---
title: "18 Test drive R Markdown"
author: "Boukara Ahmed El-Hachemi"
date: "03/07/2020"
output: 
  github_document: default
---

the official R Markdown documentation: [http://rmarkdown.rstudio.com](http://rmarkdown.rstudio.com/)  

## 1 Hello World

Launch RStudio in a Project that is a Git repo that is connected to a GitHub repo.  

Do this: *File > New File > R Markdown …*

* Give it an informative title. This will appear in the document but does not necessarily have anything to do with the file’s name. But the title and filename should be related! Why confuse yourself?   
* Accept the default Author or edit if you wish.  
* Accept the default output format of HTML.   
* Click OK.  

Save this document to a reasonable filename and location. The filename should end in `.Rmd` or `.rmd`. Save in the top-level of this RStudio project and Git repository, that is also current working directory. Trust me on this and do this for a while.

You might want to commit here. So you can see what’s about to change …

Click on “Knit HTML” or do File > Knit Document. RStudio should display a preview of the resulting HTML. Also look at the file browser. You should see the R Markdown document, i.e. `foo.Rmd` AND the resulting HTML `foo.html`.

Congratulations, you’ve just made your first reproducible report with R Markdown.

You might want to commit here.   

## 2 Push to GitHub

Push the current state to GitHub.   

Go visit it in the browser.   

Do you see the new files? An R Markdown document and the associated HTML? Visit both in the browser. Verify this:   

* Rmd is quite readable. But the output is obviously not there.  
* HTML is ugly.  

## 3 Output format

The magical process that turns your R Markdown to HTML is like so: `foo.Rmd --> foo.md --> foo.html`. Note the intermediate markdown, `foo.md`. By default RStudio discards this, but you might want to hold on to that markdown.    

In many cases, you *only want the markdown*. In that case, we switch the output format to `github_document`.  If you still want the HTML *but also the intermediate markdown*, click on the “gear” in the top bar of the source editor, near the “Knit HTML” button. Select “Output options” and go to the Advanced tab and check “Keep markdown source file.” Your YAML should now look more like this:  

```{r YAML, eval=FALSE}
    ---
    title: "Something fascinating"
    author: "Jenny Bryan"
    date: "2020-06-16"
    output:
      html_document:
        keep_md: true
    ---
```

You should have gained the line keep_md: true. You can also simply edit the file yourself to achieve this.   

```{r YAML, eval=FALSE}
    ---
    title: "Something fascinating"
    author: "Jenny Bryan"
    date: "2020-06-16"
    output: github_document
    ---
```

Save!    

You might want to commit here.    

Render via “Knit HTML” button.   

Now revisit the file browser. In addition to `foo.Rmd`, you should now see `foo.md`. If there are R chunks that make figures, the usage of markdown output formats will also cause those figure files to be left behind in a sensibly named sub-directory, `foo_files`.    

If you commit and push `foo.md` and everything inside `foo_files`, then anyone with permission to view your GitHub repo can see a decent-looking version of your report.   

If your output format is `html_document`, you should still see `foo.html`. If your output format is `github_document` and you see `foo.html`, that’s leftover from earlier experiments. Delete that. It will only confuse you later.    

You might want to commit here.   

## 4 Push to GitHub

## 5 Put your stamp on it

## 6 Develop your report

## 7 Publish your report

Creating, committing, and pushing markdown is a very functional, lighweight publishing strategy. Use `output: github_document` or `keep_md: true` if output is `html_document`. In both cases, it is critical to also commit and push everything inside foo_files. Now people can visit and consume your work like any other webpage.    

## 8 HTML on GitHub

there are ways to get a preview: such as [http://htmlpreview.github.io](http://htmlpreview.github.io/)   

## 9 Troubleshooting

Make sure RStudio and the rmarkdown package (and its dependencies) are up-to-date. Details on the system used to render this document and how to check your setup:    

* rmarkdown version 2.2. Use `packageVersion("rmarkdown")` to check yours.  
* R version 4.0.0 (2020-04-24). Use `R.version.string` to check yours.  
* RStudio IDE 1.2.1555. Use RStudio > About RStudio or `RStudio.Version()$version` to check yours.  

**Get rid of your `.Rprofile`**, at least temporarily. I have found that a “mature” `.Rprofile` that has accumulated haphazardly over the years can cause trouble. Specifically, if you’ve got anything in there relating to `knitr`, `markdown`, `rmarkdown` and RStudio itself, it may be preventing the installation or usage of the most recent goodies. Comment the whole file out or rename it something else and relaunch or even re-install RStudio.  

**“I have ignored your advice and dumped a bunch of code in at once. Now my Rmd does not render.”** If you can’t figure out what’s wrong by reading the error messages, pick one:   

* Back out of these changes, get back to a functional state (possibly with no code), and restore them gradually. Run your code interactively to make sure it works. Render the entire document frequently. Commit after each successful addition! When you re-introduce the broken code, now it will be part of a small change and the root problem will be much easier to pinpoint and fix.    
* Tell knitr to soldier on, even in the presence of errors. Some problems are easier to diagnose if you can execute specific R statements during rendering and leave more evidence behind for forensic examination.    
  * Insert this chunk near the top of your `.Rmd` document:   
```{r setup, include = FALSE, cache = FALSE}
knitr::opts_chunk$set(error = TRUE)
```

  * If it’s undesirable to globally accept errors, you can still do this for a specific chunk like so:
```{r wing-and-a-prayer, error = TRUE}
## your sketchy code goes here ;)
```

* Adapt the “git bisect” strategy:
  * Put `knitr::knit_exit()` somewhere early in your `.Rmd` document, either in inline R code or in a chunk. Keep moving it earlier until things work. Now move it down in the document. Eventually you’ll be able to narrow down the location of your broken code well enough to find the line(s) and fix it.

**Check your working directory**. It’s going to break your heart as you learn how often your mistakes are really mundane and basic. Ask me how I know. When things go wrong consider:  

  * What is the working directory?
  * Is that file I want to read/write actually where I think it is?  
  
Drop these commands into R chunks to check the above:  

  * `getwd()` will display working directory at **run time**. If you monkeyed around with working directory with, e.g., the mouse, maybe it’s set to one place for your interactive development and another when “Knit HTML” takes over?
  * `list.files()` will list the files in working directory. Is the file you want even there?    
  
**Don’t try to change working directory within an R Markdown document**. Just don’t. See [knitr FAQ #5](https://yihui.name/knitr/faq/). That is all.  

**Don’t be in a hurry to create a complicated sub-directory structure**. RStudio/`knitr`/`rmarkdown` (which bring you the “Knit HTML” button) are rather opinionated about the working directory being set to the .Rmd file’s location and about all files living together in one big happy directory. This can all be worked around. For example, I [recommend the here package](https://github.com/jennybc/here_here#readme) for building file paths, once you require sub-directories. But don’t do this until you really need it.




